---
title: "r4ds_communicate_exercises_28"
author: "Hayden Roche"
date: "November 2, 2017"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tibble)
library(stringr)
library(dplyr)
```

#Question 28.2.1.1

```{r}
ggplot(mpg, aes(displ, cty)) + 
  geom_point(aes(color = manufacturer)) +
  labs(title = "City Fuel Efficiency by Manufacturer",
       subtitle = "Chervolet's skew far to the right range of engine displacement",
       caption = "Data from fueleconomy.gov",
       x = "Engine Displacement",
       y = "City mpg",
       color = "Manufacturer")
```

#Question 28.2.1.2

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size",
       subtitle = "Two seaters (sports cars) are an exception because of their light weight",
       caption = "Data from fueleconomy.gov")
```

#Question 28.2.1.3
Done using an old swirl plot 


#Question 28.3.1.1

```{r}
label <- tribble(
  ~price, ~carat, ~label, ~vjust, ~hjust,
  Inf,  Inf,    "Top right", "top", "right",
  Inf, -Inf,    "Bottom right", "bottom", "right",
  -Inf,  Inf,    "Top left", "top", "left",
  -Inf, -Inf,    "Bottom left", "bottom", "left")

ggplot(diamonds, aes(price, carat)) +
  geom_point() +
  geom_text(aes(label = label, vjust = vjust, hjust = hjust), data = label)
```

#Question 28.3.1.2

```{r}
ggplot(diamonds, aes(price, carat)) +
  geom_point() +
  annotate("text", x = Inf, y = Inf, label = "The price of diamonds increases \nwith the overall carat size", vjust = "top", hjust = "right")

```

#Question 28.3.1.3

```{r}
label <- tibble(
  price = Inf,
  carat = Inf,
  label = "The price of diamonds increases \nwith the overall carat size")

ggplot(diamonds, aes(price, carat)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "left") +
  facet_wrap(~cut)
```

#Question 28.3.1.4

You can utilize label.padding which changes the amount of padding around label. The default is 0.25 lines. Also, label.r creates a radius of rounded corners. It defaults to 0.15 lines

#Question 28.3.1.5

 - Angle
 - Length
 - Ends
 - Type: this will change whether the arrow head is a closed or open triangle
 
#Question 28.4.4.1
It's not overriding the default scale, because geom_hex() doesn't accept the colour argument. Geom_hex() works with the fill() argument 

```{r}
df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_fill_gradient(low = "white", high = "red") +
  coord_fixed()
```

#Question 28.4.4.2

It appears from the examples that the first argument to every scale function is an argument sued to control name and labels. Therefore, it is very similar to the labs() function

#Question 28.4.4.3

```{r}
presidential %>%
  mutate(Number = 33 + row_number(),
         Name_Number = str_c(name, " (", Number, ")"),
         Name_Number = factor(Name_Number, levels = Name_Number)) %>%
  ggplot(aes(start, Name_Number, colour = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = Name_Number)) +
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue")) 
```

#Question 28.4.4.4

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(colour = cut), alpha = 1/20) +
  theme(legend.position = "top") +
  guides(colour = guide_legend(nrow = 1, override.aes = list(alpha = 1/2)))
```



