---
title: "Reproducible Research - Project 2"
author: "Hayden Roche"
date: "December 19, 2017"
output: html_document
---

#Data Processing

Using a download.file function I was able to pull the zip file right into R Studio and read the csv in through a local storage file. The next step was implementing gsub to clean up the PROMDMG column because there were a lot of inconsistencies with the naming conventions. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(dplyr)
library(RCurl)
library(ggplot2)

download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2", "storm.csv")
Storm_raw <- read.csv("storm.csv")

Clean_Storm_data <- 
  Storm_raw[,c("BGN_DATE","EVTYPE", "FATALITIES", "INJURIES", "PROPDMG", "PROPDMGEXP", 'CROPDMG', 'CROPDMGEXP')]
Clean_Storm_data$BGN_DATE <- as.Date(Clean_Storm_data$BGN_DATE, format = "%m/%d/%Y")
Clean_Storm_data$TotalDamage <- paste(Clean_Storm_data$PROPDMG,"-", Clean_Storm_data$PROPDMGEXP)
Clean_Storm_data$EVTYPE <- as.character(Clean_Storm_data$EVTYPE)
Clean_Storm_data$EVTYPE <- 
  gsub("^HEAT", "EXCESSIVE HEAT", Clean_Storm_data$EVTYPE)
Clean_Storm_data$EVTYPE <- 
  gsub("^TSTM WIND$", "THUNDERSTORM WIND", Clean_Storm_data$EVTYPE)
Clean_Storm_data$EVTYPE <- 
  gsub("^THUNDERSTORM WIND$", "THUNDERSTORM WINDS", Clean_Storm_data$EVTYPE)
Clean_Storm_data$EVTYPE <- 
  gsub("^RIP CURRENTS$", "RIP CURRENT", Clean_Storm_data$EVTYPE)

```


#Synopsis
This analysis is looking into fatalities, injuries and property damages from various natural disasters. These natural disasters have been recorded over from 1950-2011. The overall findings are quite interesting: torandoes, excessive heat and flash floods are the deadliest of natural disasters. Over the 60 years of data collection, tornadoes accounted for 5.6K fatalities.

The story is similar when we look at injuries, but the magnitude increases significantly. Tornadoes led the injury count by miles, with a total 91.3K and then excessive heat followed with 8.6K injuries. An interesting next step would be creating a weighted damage measure that factored in both deaths and injuries. This would then take a nice consistent measure for damages. 

```{r cars}
str(Clean_Storm_data)
head(Clean_Storm_data)
```

#Results
##Fatalities: Manipulating the data to the top 5 most dangerous natural events type
```{r fatalities, echo=TRUE}
Fatalities_DF <- Clean_Storm_data %>%
  mutate(Year = year(BGN_DATE)) %>%
  group_by(Year, EVTYPE) %>%
  summarise(All_Fatalities = sum(FATALITIES)) %>%
  filter(EVTYPE == 'EXCESSIVE HEAT' | EVTYPE == 'TORNADO' | EVTYPE == 'LIGHTNING' | EVTYPE == 'FLASH FLOOD' | EVTYPE == 'RIP CURRENT' | EVTYPE == 'FLOOD' | EVTYPE == 'HIGH WIND' | EVTYPE == 'AVALANCHE')

m <- ggplot(data = Fatalities_DF, aes(x = Year, y = All_Fatalities)) +
  geom_point() +
  stat_smooth() 
  m + facet_wrap( ~ EVTYPE, ncol=4, scales="free")
print(m)
```

##Fatalities: Manipulating the data to the top 5 most dangerous natural events type
```{r fatalities2, echo=TRUE}
Fatalities_Total <- Clean_Storm_data %>%
  mutate(Year = year(BGN_DATE)) %>%
  group_by(EVTYPE) %>%
  summarise(All_Fatalities = sum(FATALITIES)) %>%
  as.data.frame() %>%
  arrange(desc(All_Fatalities)) %>%
  filter(EVTYPE == 'EXCESSIVE HEAT' | EVTYPE == 'TORNADO' | EVTYPE == 'LIGHTNING' | EVTYPE == 'FLASH FLOOD' | EVTYPE == 'RIP CURRENT' | EVTYPE == 'FLOOD' | EVTYPE == 'HIGH WIND' | EVTYPE == 'AVALANCHE')
View(Fatalities_Total)

b <- ggplot(Fatalities_Total, aes(x = EVTYPE, y = All_Fatalities, fill = EVTYPE))
b + geom_bar(stat = "identity")
```

##Injuries: Manipulating the data to the top 8 most dangerous natural events type
```{r injuries, echo=TRUE}
Injuries_DF <- Clean_Storm_data %>%
  mutate(Year = year(BGN_DATE)) %>%
  group_by(Year, EVTYPE) %>%
  summarise(All_Injuries = sum(INJURIES)) %>%
  filter(EVTYPE == 'EXCESSIVE HEAT' | EVTYPE == 'TORNADO' | EVTYPE == 'LIGHTNING' | EVTYPE == 'FLASH FLOOD' | EVTYPE == 'RIP CURRENT' | EVTYPE == 'FLOOD' | EVTYPE == 'HIGH WIND' | EVTYPE == 'AVALANCHE')

m <- ggplot(data = Injuries_DF, aes(x = Year, y = All_Injuries)) +
  geom_point() +
  stat_smooth() 
m + facet_wrap( ~ EVTYPE, ncol=4, scales="free")
print(m)
```

```{r injuries2, echo=TRUE}
Injuries_Total <- Clean_Storm_data %>%
  mutate(Year = year(BGN_DATE)) %>%
  group_by(EVTYPE) %>%
  summarise(All_Injuries = sum(INJURIES)) %>%
  as.data.frame() %>%
  arrange(desc(All_Injuries)) %>%
  filter(EVTYPE == 'EXCESSIVE HEAT' | EVTYPE == 'TORNADO' | EVTYPE == 'LIGHTNING' | EVTYPE == 'FLASH FLOOD' | EVTYPE == 'RIP CURRENT' | EVTYPE == 'FLOOD' | EVTYPE == 'ICE STORM' | EVTYPE == 'HAIL')
View(Injuries_Total)

b <- ggplot(Injuries_Total, aes(x = EVTYPE, y = All_Injuries, fill = EVTYPE))
b + geom_bar(stat = "identity")
```

##Property Damage
```{r property, echo=TRUE}
Property <- Clean_Storm_data %>%
  mutate(Year = year(BGN_DATE)) %>%
  group_by(EVTYPE) %>%
  summarise(All_Damage = sum(PROPDMG)) %>%
  as.data.frame() %>%
  arrange(desc(All_Damage)) %>%
  filter(EVTYPE == 'WINTER STORM' | EVTYPE == 'TORNADO' | EVTYPE == 'LIGHTNING' | EVTYPE == 'FLASH FLOOD' | EVTYPE == 'THUNDERSTORM WINDS' | EVTYPE == 'FLOOD' | EVTYPE == 'HIGH WIND' | EVTYPE == 'HAIL')
View(Property)

d <- ggplot(Property, aes(x = EVTYPE, y = All_Damage, fill = EVTYPE))
d + geom_bar(stat = "identity")
```
